CudeleFS: Programmable Consistency and \\Durability in a Global Namespace

### 2017-04-04 paper/sections/3_posix_overheads.tex
paper: jeff/carlos intro comments, fix figs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% - sequential IO, trim redundant operations
\noindent\textbf{CephFS Design}: a journal of metadata updates that streams into
the resilient object store. Similar to LFS~\cite{rosenblum:acm1992-LFS} and
WAFL~\cite{hitz:wtec1994-WAFL} the metadata journal is designed to grow to large which
ensures (1) sequential writes into the object store and (2) the ability for daemons to
trim redundant or irrelevent journal entries.  The journal is striped over
objects where multiple journal updates can reside on the same object. There are
two tunables for controlling the journal: the segment size and the number of
parallel segments that can be written in parallel. Unless the journal saturates 
memory or CPU resources, larger values for these tunables results in better
performance.

@@

% benefits PROBLEM -- IS THIS THE METADATA PROTOCOL OR JUST THE OVERLOADEDNESS?
The benefits of caching the directory inode when creating files is shown in
Figure~\ref{fig:overhead-b}.  If only one client is creating files in a
directory (``isolated" curve on \(y1\) axis) then that client can lookup the
existence of new files locally before issuing a create request to the metadata
server. If another client starts creating files in the same directory
(``interfere" curve on \(y1\) axis) then the directory inode transitions out of
read caching and the first client must send \texttt{lookup()}s to the metadata
server (``interfere" curve on \(y2\) axis). These extra requests increase the
throughput of the ``interfere" curve because the metadata server can handle the
extra load but the overall performance decreases.  This degradation is shown in
Figure~\ref{fig:overhead-c}, where we scale the number of clients and show
increased variability. The performance is compared to a single
isolated client and the \(y\) axis the standard deviation of the runtime of
all clients.  For the ``interfere" bars, each client creates files in private
directories and at 30 seconds we launch another process that creates files in
those directories.  For less than 7 clients, the runtime and deviation is worse
when clients interfere. At 7 clients, the metadata server is overloaded so the
directory caching in the ``isolated" bars has no benefit.\\

>>>>

% - sequential IO, trim redundant operations
\noindent\textbf{CephFS Design}: a journal of metadata updates that streams into
the resilient object store. Similar to LFS~\cite{rosenblum:acm1992-LFS} and
WAFL~\cite{hitz:wtec1994-WAFL} the metadata journal is designed to grow to large which
ensures (1) sequential writes into the object store and (2) the ability for daemons to
trim redundant or irrelevant journal entries.  The journal is striped over
objects where multiple journal updates can reside on the same object. There are
two tunables for controlling the journal: the segment size and the number of
parallel segments that can be written in parallel. Unless the journal saturates 
memory or CPU resources, larger values for these tunables results in better
performance.

@@

% benefits PROBLEM -- IS THIS THE METADATA PROTOCOL OR JUST THE OVERLOADEDNESS?
The benefits of caching the directory inode when creating files is shown in
Figure~\ref{fig:overhead-b}.  If only one client is creating files in a
directory (``isolated" curve on \(y1\) axis) then that client can lookup the
existence of new files locally before issuing a create request to the metadata
server. If another client starts creating files in the same directory
(``interfere" curve on \(y1\) axis) then the directory inode transitions out of
read caching and the first client must send \texttt{lookup()}s to the metadata
server (``interfere" curve on \(y2\) axis). These extra requests increase the
throughput of the ``interfere" curve because the metadata server can handle the
extra load but performance suffers.  This degradation is shown in
Figure~\ref{fig:overhead-c}, where we scale the number of clients and show
increased variability. The results are normalized to a single
isolated client and the \(y\) axis is the standard deviation of the runtime of
all clients.  For the ``interfere" bars, each client creates files in private
directories and at 30 seconds we launch another process that creates files in
those directories.  For less than 7 clients, the runtime and deviation is worse
when clients interfere. At 7 clients, the metadata server is overloaded so the
directory caching in the ``isolated" bars has no benefit.\\

### 2017-04-04 paper/sections/6_results.tex
paper: jeff/carlos intro comments, fix figs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Figure~\ref{fig:slowdown-mechanisms} shows the runtime of the CudeleFS
mechanisms, normalized to the time it takes to write \(n\) file create updates
to the client's in-memory journal ({\it i.e.} the ``Append Client Journal""
mechanism). Bars above \(10^0\) are slower than the ``Append Client Journal""
mechanism and bars below are faster.  ``Stream" is an approximation of the
overhead and is calculated by subtracting the runtime of the job with the
journal turned off from the runtime with the journal turned on.  The largest
workoad we tested is 100K file creates in the same directory.  This is the
maximum recommended size of a directory in CephFS; preliminary experiemnts with
larger directory sizes show memory problems.

@@

% TODO: why is apply so slow.  
% apply: no CephFS changes, pulls/pushes same RADOS obj.
% v_apply vs. apply/persist: communicating through RADOS
\subsubsection{Overhead of ``Nonvolatile Apply"} The cost of ``Nonvolatile
apply" is much larger than all the other mechanisms.  That mechanism was not
implemented as part of CudeleFS -- it was a debugging and recovery tool packaged
with CephFS. It works by iterating over the updates in the journal and pulling
all objects that {\it may} be affected by the update.  This means that two
objects are repeatedly pulled, updated, and pushed: the object that houses the
experiment directory and the object that contains the root directory ({\it
i.e.} \texttt{/}).  The cost of communcating through the object store is shown
by comparing the runtime of ``Volatile apply" + ``Global persist" to
``Nonvolatile Apply". These two operations end up with the same final metadata
state but using ``Nonvolatile Apply" is clearly inferior.

@@

\noindent\textbf{Takeaway}: the CudeleFS mechansisms have overheads and costs
that can differ {\it by order of magnitude}. CudeleFS gives users the ability
to compose these mechanisms based on their application's correctness
requirements and performance goals.

@@

Figure~\ref{fig:slowdown-strong-weak} shows the runtimes of systems employing
weak and strong consistency, normalized to the runtime of the ``Append Client Journal""
mechanism (again, just creating files in the client's in-memory journal).  We
use the following compositions from the mechanisms in
Table~\ref{table:spectrum}:  Strong Consistency = ``RPCs" + ``Stream"; Strong
Consistency, No Fault Tolerance = ``RPCs"; Weak Consistency = ``Append Client Journal"s" +
``Local Persist"; and Weak Consistency, No Fault Tolerance = ``Append Client Journal"s" +
``Local Persist" + ``Volatile Apply".

@@

% POSIX (no stream) vs. POSIX: cost of durability < consistency
\subsubsection{Durability \(<<\) Consistency} The \(1.15\times\) overhead of
Strong Consistency compared to Strong Consistency, No Fault Tolerance for
100K files is neglible. It suggests that the overhead of consistency is much
larger than the overhead of durability. This conclusion should be stronger as
we scale the number of files because the cost of streaming the journal into the
object store is constant. We omit the same analysis for Weak Consistency
because the runtimes are so short that the normalized slowdowns are misleading.

@@

\begin{listing}[tb]
\begin{minted}[frame=single,
               framesep=3mm,
               xleftmargin=21pt,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "RPCs"
    "durability": "stream"
}
\end{minted}
\caption{Existing CephFS on CudeleFS with blocking.}
\label{src:block}
\end{listing}

\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/slowdown-allow-block.png}
\caption{
[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/viz-blockallow.ipynb}{source}]
Using the ``allow" and ``block" API, users can isolate directories from
interferring clients. Performance with blocking turned is the same as
``isolated" from Figure~\ref{fig:interfere-b}.
\label{fig:slowdown-allow-block}}
\end{figure}

% setup: 2cs sep dirs, 2cs DN, 1c malicious write
Next we show how CudeleFS can be programmed to block interferring clients. We
use the same problematic workload from Figure~\ref{fig:interfere-b}, where
clients write to their own private directories and another client interferes with a stream of creates at
30 seconds.  We also have another client write to a decoupled namespace and
merge its updates at 90 seconds.  We equip each client directory with the
configuration in Listing~\ref{src:block}.  Each directory functions with RPCs
and stream enabled -- which is the default implementation of CephFS. The only
difference is that we enable blocking on each subtree so while the tree behaves
like CephFS, all interferring operations will be returned with \texttt{-EBUSY}.
Note that IndexFS does a similar operation with leases except clients block.

% results
To show the benefits of this isolations, our results in
Figure~\ref{fig:slowdown-allow-block} are plotted alongside the bars from
Figure~\ref{fig:interfere-b}. Because ``allow"/``interfere and
``block"/``isolated" have the same performance we draw the following three
conclusions: (1) clients that use the API to block interferring clients  get
the same performance as isolated clients, (2) there is a neglible effect on
performance for the extra work the metadata server does to return
\texttt{-EBUSY}, and (3) merging updates from the decoupled client has a
neglible effect on performance.\\

>>>>

Figure~\ref{fig:slowdown-mechanisms} shows the runtime of the CudeleFS
mechanisms, normalized to the time it takes to write \(n\) file create updates
to the client's in-memory journal ({\it i.e.} the ``Append Client Journal"
mechanism). Bars above \(10^0\) are slower than the ``Append Client Journal"
mechanism and bars below are faster.  ``Stream" is an approximation of the
overhead and is calculated by subtracting the runtime of the job with the
journal turned off from the runtime with the journal turned on.  The largest
workload we tested is 100K file creates in the same directory.  This is the
maximum recommended size of a directory in CephFS; preliminary experiments with
larger directory sizes show memory problems.

@@

% TODO: why is apply so slow.  
% apply: no CephFS changes, pulls/pushes same RADOS obj.
% v_apply vs. apply/persist: communicating through RADOS
\subsubsection{Overhead of ``Nonvolatile Apply"} The cost of ``Nonvolatile
apply" is much larger than all the other mechanisms.  That mechanism was not
implemented as part of CudeleFS -- it was a debugging and recovery tool packaged
with CephFS. It works by iterating over the updates in the journal and pulling
all objects that {\it may} be affected by the update.  This means that two
objects are repeatedly pulled, updated, and pushed: the object that houses the
experiment directory and the object that contains the root directory ({\it
i.e.} \texttt{/}).  The cost of communicating through the object store is shown
by comparing the runtime of ``Volatile apply" + ``Global persist" to
``Nonvolatile Apply". These two operations end up with the same final metadata
state but using ``Nonvolatile Apply" is clearly inferior.

@@

\noindent\textbf{Takeaway}: the CudeleFS mechanisms have overheads and costs
that can differ {\it by order of magnitude}. CudeleFS gives users the ability
to compose these mechanisms based on their application's correctness
requirements and performance goals.

@@

Figure~\ref{fig:slowdown-strong-weak} shows the runtime of systems employing
weak and strong consistency, normalized to the runtime of the ``Append Client Journal"
mechanism (again, just creating files in the client's in-memory journal).  We
use the following compositions from the mechanisms in
Table~\ref{table:spectrum}:  Strong Consistency = ``RPCs" + ``Stream"; Strong
Consistency, No Fault Tolerance = ``RPCs"; Weak Consistency = ``Append Client Journal" +
``Local Persist"; and Weak Consistency, No Fault Tolerance = ``Append Client Journal" +
``Local Persist" + ``Volatile Apply".

@@

% POSIX (no stream) vs. POSIX: cost of durability < consistency
\subsubsection{Durability \(<<\) Consistency} The \(1.15\times\) overhead of
Strong Consistency compared to Strong Consistency, No Fault Tolerance for
100K files is negligible. It suggests that the overhead of consistency is much
larger than the overhead of durability. This conclusion should be stronger as
we scale the number of files because the cost of streaming the journal into the
object store is constant. We omit the same analysis for Weak Consistency
because the runtimes are so short that the normalized slowdowns are misleading.

@@

\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/slowdown-allow-block.png}
\caption{
[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/viz-blockallow.ipynb}{source}]
Using the ``allow" and ``block" API, users can isolate directories from
interfering clients. Performance with blocking turned is the same as
``isolated" from Figure~\ref{fig:overhead-b}.
\label{fig:slowdown-allow-block}}
\end{figure}

% setup: 2cs sep dirs, 2cs DN, 1c malicious write
Next we show how CudeleFS can be programmed to block interfering clients. We
use the same problematic workload from Figure~\ref{fig:overhead-b}, where
clients write to their own private directories and another client interferes with a stream of creates at
30 seconds.  We also have another client write to a decoupled namespace and
merge its updates at 90 seconds.  We equip each client directory with the
configuration:

\begin{listing}[tb]
\begin{minted}[xleftmargin=21pt,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "RPCs"
    "durability": "stream"
}
\end{minted}
\end{listing}

Each directory functions with ``RPCs"
and ``Stream" enabled -- which is the default implementation of CephFS. The only
difference is that we enable blocking on each subtree so while the tree behaves
like CephFS, all interfering operations will be returned with \texttt{-EBUSY}.
Note that IndexFS does a similar operation with leases except clients block.

% results
To show the benefits of this isolations, our results in
Figure~\ref{fig:slowdown-allow-block} are plotted alongside the variance bars from
Figure~\ref{fig:interfere-b}. Because ``allow"/``interfere and
``block"/``isolated" have the same variability we draw the following three
conclusions: (1) clients that use the API to block interfering clients  get
the same performance as isolated clients, (2) there is a negligible effect on
performance for the extra work the metadata server does to return
\texttt{-EBUSY}, and (3) merging updates from the decoupled client has a
negligible effect on performance.\\
